---
title: "liver transplant request 11122025"
output: html_document
date: "2025-11-12"
---

```{r,include=FALSE}
library(tidyverse)
library(tableone)
library(writexl)
library(gtsummary)
library(survival)
library(ggplot2)
library(readxl)
library(ggpubr)
library(splines)
library(xlsx)
library(lattice)
library(mice)
library(DescTools)
library(survminer)
library(finalfit)
library(tidycmprsk)
library(ggsurvfit) 
library(sqldf)
library(table1)
library(cobalt)
library(MatchIt)



final_master <- read_xlsx("/Users/to909/Desktop/Terlipressin/final_master_0508.xlsx")

final_master <- final_master %>%
  mutate(across(everything(), ~{
    x_char <- as.character(.x)
    if (all(grepl("^[-+]?[0-9]*\\.?[0-9]+$", x_char[!is.na(x_char)]))) {
      as.numeric(x_char)
    } else {
      .x  # leave unchanged
    }
  }))

final_master <- final_master %>%
  mutate(
    BMI = aki_weight / (aki_height/100)^2,
    rrt_90days=ifelse(is.na(days_to_rrt),0,ifelse(days_to_rrt<=90,1,0)),
    lt_90days=ifelse(is.na(time_to_transplant),0,ifelse(time_to_transplant<=90,1,0)),
   rrt_status=case_when((days_to_rrt -  time_to_transplant) == 0 ~ 0,
                         (days_to_rrt -  time_to_transplant) < 0 ~ 1,
                         (days_to_rrt -  time_to_transplant) > 0 ~ 2))
  


```



```{r,include=FALSE}

# create scr_day
map_days <- grep("^map_terli_day", str_sort(names(final_master)), value = TRUE)
map_day0 <- final_master %>% 
  select(subjectid,contains("map_terli_day0_time"))

df_map_day0 <- map_day0 %>%  
  pivot_longer(
    cols = starts_with("map_terli_day"), 
    names_to = c("temp", "day", "time"),    # we will capture these parts
    names_pattern = "^(map_terli)_day(\\d+)_time(\\d+)$", # regex to split
    values_to = "map"
  ) %>%
  # 2) Convert day/time to numeric so we can sort properly
  mutate(
    day = as.numeric(day),
    time = as.numeric(time)
  ) %>%
  select(-temp)  # 'temp' is just the literal "map_terli", can drop if not needed

final_map_day0 <- df_map_day0 %>% 
  group_by(subjectid, day) %>% 
   summarise(map_day0=mean(map,na.rm=T)) %>% select(subjectid,map_day0) %>% ungroup()

```

```{r,include=FALSE}

# data: a data.frame (e.g., final_master)
# prefix: the column-name prefix before the day number (e.g., "scr_terli_day")
# new_var_name: name of the new column to create (e.g., "scr_terli_discont_day")
add_by_discont_day <- function(data, prefix, new_var_name) {
  if (!is.data.frame(data)) stop("`data` must be a data.frame.")
  if (!"Discont_Day" %in% names(data)) stop("Column `Discont_Day` not found in `data`.")

  # Escape regex specials in the prefix, then build a regex to capture the trailing day number
  esc_prefix <- gsub("([\\^$.*+?(){}|\\[\\]\\\\])", "\\\\\\1", prefix)
  rx <- paste0("^", esc_prefix, "(\\d+)$")

  # Find day columns like "<prefix><day>", e.g. "scr_terli_day5"
  day_cols <- grep(rx, names(data), value = TRUE)
  if (length(day_cols) == 0) stop("No columns found matching the given `prefix` followed by a day number.")

  # Extract their day numbers and order columns by that number
  day_nums <- as.integer(sub(rx, "\\1", day_cols))
  ord <- order(day_nums)
  day_cols <- day_cols[ord]
  day_nums <- day_nums[ord]

  # Build a matrix for fast row-wise indexing
  mat <- as.matrix(data[, day_cols, drop = FALSE])

  # For each row, find which column index matches its Discont_Day
  dd <- data[["Discont_Day"]]
  col_index <- match(dd, day_nums)

  # Pull values using matrix indexing; NAs when no matching day/column
  out <- rep(NA, nrow(data))
  take <- !is.na(col_index)
  if (any(take)) {
    out[take] <- mat[cbind(which(take), col_index[take])]
  }

  # Add the new column and return the modified data
  data[[new_var_name]] <- out
  data
}
```

```{r,include=FALSE}


final_master <- final_master %>% merge(final_map_day0,by="subjectid",all.x = T)

final_master <- final_master %>% mutate(ifelse(is.na(refascites_previous_comp),0,refascites_previous_comp))


final_master <- final_master %>% mutate_at(c("hrs_responders_cat_2","sex_male"),as.factor)

final_master <- add_by_discont_day(
  data = final_master,
  prefix = "alb_terli_day",
  new_var_name = "alb_terli_discont_day"
)
final_master_listed_transplant <- final_master %>% filter(listed_transplant==1)





```

# Transplant yes = 2 no = 0, transplant_lt_status = 2 
```{r}
table(final_master_listed_transplant$transplant_lt_status)

```

# rrt_status
```{r}
table(final_master_listed_transplant$rrt_status)

```

# Death in 90 days yes = 1 no = 0, time_to_death_status_90days = 1 
```{r}
table(final_master_listed_transplant$time_to_death_status_90days)

```


# rrt_status, hrs_responders_cat_2
```{r}
table(final_master_listed_transplant$rrt_status,final_master_listed_transplant$hrs_responders_cat_2)
fisher.test(table(final_master_listed_transplant$rrt_status,final_master_listed_transplant$hrs_responders_cat_2))

```


# (listed_transplant=1) and transplanted (transplant_lt_status = 2) new$rrt_90days,new$hrs_responders_cat_2 
```{r}
new <- final_master_listed_transplant %>% filter(listed_transplant==1 & transplant_lt_status == 2)
table(new$rrt_90days,new$hrs_responders_cat_2)
chisq.test(table(new$rrt_90days,new$hrs_responders_cat_2))
```


# transplant_lt_status vs  hrs_responders_cat_2
```{r}

table(final_master_listed_transplant$transplant_lt_status,final_master_listed_transplant$hrs_responders_cat_2)
chisq.test(table(final_master_listed_transplant$transplant_lt_status,final_master_listed_transplant$hrs_responders_cat_2))
```

# transplant_lt_status = 2    transplant_slkt vs hrs_responders_cat_2
```{r}
new <- final_master_listed_transplant %>% filter(transplant_lt_status == 2)
table(new$transplant_slkt,new$hrs_responders_cat_2)
chisq.test(table(new$transplant_slkt,new$hrs_responders_cat_2))
```



# dc_days_to_death mean and sd
```{r}
mean(final_master_listed_transplant$dc_days_to_death,na.rm = T)
sd(final_master_listed_transplant$dc_days_to_death,na.rm = T)
summary(final_master_listed_transplant$dc_days_to_death)
```


# remove time_to_death_status_90days=0
```{r}
final_master_listed_transplant_alive <- final_master_listed_transplant %>% filter(time_to_death_status_90days!=0)
mean(final_master_listed_transplant_alive$dc_days_to_death,na.rm = T)
sd(final_master_listed_transplant_alive$dc_days_to_death,na.rm = T)
summary(final_master_listed_transplant_alive$dc_days_to_death)
```



# time_to_transplant_90days mean and sd
```{r}
mean(final_master_listed_transplant$time_to_transplant_90days,na.rm = T)
sd(final_master_listed_transplant$time_to_transplant_90days,na.rm = T)
summary(final_master_listed_transplant$time_to_transplant_90days)
```

# dc_days_to_death mean and sd
```{r}
mean(final_master_listed_transplant$dc_days_to_death,na.rm = T)
sd(final_master_listed_transplant$dc_days_to_death,na.rm = T)
summary(final_master_listed_transplant$dc_days_to_death)
```


# Competing risk model outcome transplant=2, death=1 
```{r}
table(final_master_listed_transplant$time_to_death_status_90days_cmp)

```

# Competing risk model outcome 2 competing event 1 covariates:hrs_responders_cat_2+age+sex_male+admit_meld_3
```{r}
final_master_listed_transplant$time_to_death_status_90days_cmp <- factor(final_master_listed_transplant$time_to_death_status_90days_cmp)


tidycmprsk::crr(Surv(time_to_death_90days_cmp,time_to_death_status_90days_cmp)~ hrs_responders_cat_2+age+sex_male+admit_meld_3,failcode=2,cencode=0, data = final_master_listed_transplant) %>% tbl_regression(exp = TRUE) %>% add_n()%>%
  bold_p(0.05) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() 

```

# logistic regression outcome:hrs_responders_cat_2  covariates:age+sex_male+admit_meld_3+map_day0+refascites_previous_comp
```{r}

glm(hrs_responders_cat_2~age+sex_male+admit_meld_3+map_day0+refascites_previous_comp, data = final_master_listed_transplant, family = binomial) %>% tbl_regression(exp = TRUE,pvalue_fun = function(x) style_pvalue(x, digits = 3))

```