---
title: "Liver Transplant — Analysis"
author: "Tianqi Ouyang"
date: last-modified
format:
  html:
    echo: true
    code-fold: true
    toc: true
    toc-depth: 3
---

# Scope

Extended follow-up (EFU) of the HARMONY terlipressin cohort. The analytic
cohort is every patient in the locked master dataset
(`final_master_01282026.xlsx`, n = 243) who has an EFU record in the
site-submitted files under
`Extended Data Collection/Finalized Files/09SEP2026/` (13 site CSVs,
222 rows).

| Table | Cohort | Column (stratifier) | Rows |
|:--|:--|:--|:--|
| 1 — Demographics | EFU patients | `efu_listingstatus` (listed for LT at any point) | Baseline demographics, prior complications, admission / day-0 labs & scores, treatment characteristics, HRS response |
| 2 — EFU summary | EFU patients | `listed_transplant` (index-hospitalization listing, shared pipeline) | All EFU outcome fields |
| 3 — EFU summary | EFU patients | `efu_listingstatus` | All EFU outcome fields except the stratifier |

# Legacy Code Review — Errors Found

The previous analyses in `Code/*.Rmd` were reviewed before this file was
written. The following problems were found; none of them are carried
forward here.

| # | File | Error | Consequence | Handling here |
|:--|:--|:--|:--|:--|
| 1 | `liver-transplant-0922.Rmd` | `rrt_90days = ifelse(is.na(days_to_rrt) & days_to_rrt <= 90, 1, 0)` — `is.na(x) & x <= 90` can never be `TRUE` | `rrt_90days` was never 1. Same bug in `lt_90days`, which additionally compared `days_to_rrt` instead of `days_to_lt` | Fixed in the later `liver request 11122025.Rmd`; not needed here (EFU has its own RRT / LT fields) |
| 2 | `liver-transplant-0922.Rmd` | `albvolume_before_terli` listed in `num_vars` but not in `all_vars` | Silently dropped from every Table 1 | Included in `all_vars` |
| 3 | `liver-transplant-0922.Rmd` | "Table 4" chunk is an exact duplicate of "Table 3" (same strata, same output path) | Same CSV written twice | Removed |
| 4 | `liver request 11122025.Rmd` | `mutate(ifelse(is.na(refascites_previous_comp), 0, refascites_previous_comp))` has no `name =` | Creates a column literally named `ifelse(...)`; the intended NA → 0 recode never reached `refascites_previous_comp`, so the logistic model silently dropped those rows | Not applicable here (no recode requested); flagged |
| 5 | both `.Rmd` | `alb_terli_discont_day` was built from the `alb_terli_dayN` fields, which are **serum albumin (g/dL)**, not albumin *given* | The "albumin during therapy" row in the old Table 1 was a lab value, not a dose | Albumin given is `albumintotal_terli_dayN` (grams/day); the shared pipeline sums days 0–13 into `total_albumin`, used here for the requested `albumintotal_terli_dayX` row. No albumin is recorded after `Discont_Day`, so the sum equals "total albumin during therapy" |
| 6 | both `.Rmd` | Read the superseded `/Users/to909/Desktop/Terlipressin/final_master_0508.xlsx` | Predates the cleaned master | Shared pipeline → `final_master_01282026.xlsx` |
| 7 | shared pipeline | The master xlsx already contains `mean_map_day0`; the pipeline merges a second copy, producing `mean_map_day0.x` / `.y` (no `mean_map_day0` column) | Any sub-project referencing `mean_map_day0` / `map_day0` silently loses that row from its Table 1 | `map_day0_avg` is computed explicitly in this file |
| 8 | master data | `map_terli_day0_time0..3` contain `0` values (25 / 4 / 2 / 2 readings) — placeholders for "not measured", not real pressures | Averaging them in biases the day-0 MAP downward | `0` → `NA` before averaging (same convention as the Xing sub-project) |
| 9 | EFU data | Yale IDs are `YAL-00N` in the EFU file but `Yale-N` in the master | All 5 Yale patients failed to join | Join on a normalized key: upper-case prefix (`YAL` → `YALE`) + integer number |
| 10 | EFU data | One CCF row (site file has 41 rows; master has 40 CCF patients) has an ID that is not in the master at all | 222 EFU rows → **221** analyzable patients | Reported below; cohort n is computed, not hard-coded |
| 11 | EFU data | No CSF file was submitted | The 22 CSF patients in the master have no EFU data | Reported below |
| 12 | EFU data | One `efu_lttbili` value is a below-detection string (`<x.x`) | `as.numeric()` → `NA` | Parsed with `readr::parse_number()` (uses the limit value) |
| 13 | shared pipeline | `create_table_one()` renders the HTML table with `overall = "Overall"`, and `table1` passes that Overall column into the `pvalue()` helper as a **third group** | Every HTML p-value is computed on (group 0, group 1, Overall) — e.g. `sex_male` p = 0.239 instead of 0.121 — and 2-group continuous variables get a Kruskal-Wallis instead of Mann-Whitney. The tableone CSVs are unaffected | `pvalue()` is redefined below to drop the `overall` element before testing, so the HTML and CSV p-values agree |

```{r setup}
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(readxl)
library(tableone)
library(table1)
library(htmltools)

# Load `master` (n = 243) and the shared helpers. `output = tempfile()` keeps
# the purl()ed script out of the working directory.
source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  output = tempfile(fileext = ".R"),
  quiet  = TRUE
))

# --- Fix for shared-pipeline bug #13 -----------------------------------------
# table1::table1(..., overall = "Overall", extra.col = list("P-value" = pvalue))
# passes the Overall column into `pvalue()` as an extra stratum. Drop it so
# the test compares only the real strata (matches the tableone CSV).
pvalue <- function(x, ...) {
  x <- x[tolower(names(x)) != "overall"]
  y <- unlist(x)
  g <- factor(rep(1:length(x), times = sapply(x, length)))
  if (is.numeric(y)) {
    n_groups <- nlevels(droplevels(g))
    if (n_groups == 2) {
      p <- tryCatch(wilcox.test(y ~ g)$p.value, error = function(e) NA)
      test_used <<- c(test_used, "wilcox")
    } else {
      p <- tryCatch(kruskal.test(y ~ g)$p.value, error = function(e) NA)
      test_used <<- c(test_used, "kruskal")
    }
  } else {
    tbl <- table(y, g)
    if (nrow(tbl) < 2 || ncol(tbl) < 2) {
      p <- NA
      test_used <<- c(test_used, "none")
    } else if (any(tbl < 5)) {
      p <- tryCatch(fisher.test(tbl)$p.value, error = function(e) NA)
      test_used <<- c(test_used, "fisher")
    } else {
      p <- tryCatch(chisq.test(tbl)$p.value, error = function(e) NA)
      test_used <<- c(test_used, "chisq")
    }
  }
  pval <- if (is.na(p)) "" else sub("<", "&lt;", format.pval(p, digits = 3, eps = 0.001))
  c("", pval)
}
# `source()` defined the pipeline functions in the global environment, and
# knitr evaluates this chunk in a child environment, so `create_table_one()`
# would still find the pipeline's `pvalue`. Install the fixed version there.
assign("pvalue", pvalue, envir = globalenv())

dir.create("tables", showWarnings = FALSE)
```

# EFU Data Import

Each site returned one CSV. All files share the same 42 columns
(`subjectid` + 41 EFU fields; see the *EFU Variables* page). The files
contain no dates — only REDCap-calculated day counts from the index
admission — so nothing patient-identifying is loaded beyond the study ID
used for the join.

```{r efu-import}
#| code-summary: "Show EFU import and ID normalization"
efu_dir <- "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Extended Data Collection/Finalized Files/09SEP2026"

# Study IDs differ in formatting between REDCap exports (e.g. `YAL-003` in the
# EFU file vs `Yale-3` in the master). Normalize both sides to
# `<UPPERCASE PREFIX>-<integer>` before joining.
norm_id <- function(x) {
  x      <- str_trim(x)
  prefix <- toupper(str_extract(x, "^[A-Za-z]+"))
  prefix <- if_else(prefix == "YAL", "YALE", prefix)
  number <- as.integer(str_extract(x, "[0-9]+$"))
  paste0(prefix, "-", number)
}

# Field types from the EFU data dictionary
efu_yesno <- c(
  "codestatus", "efu_readmit", "efu_readmit_cirrhosiscomp", "efu_readmitaki",
  "efu_status", "efu_rrt", "efu_listingstatus", "efu_listing", "efu_lt",
  "efu_slkt", "efu_kal", "efu_ltrrt", "efu_90_rrt", "efu_90lt_rrt",
  "efu_180_rrt", "efu_180lt_rrt", "efu_1yr_rrt", "efu_1yrlt_rrt"
)
efu_numeric <- c(
  "efu_daysreadmit", "efu_readmitscr", "efu_dayslastencounter", "efu_daysdod",
  "efu_daysrrtstart", "efu_dayslisting", "efu_dayswlremoval", "efu_dayslt",
  "efu_dayskal", "efu_ltna", "efu_lttbili", "efu_ltscr", "efu_lt_albumin",
  "efu_lt_inr", "efu_90_scr", "efu_90_readmit", "efu_90lt_scr", "efu_180_scr",
  "efu_180_readmit", "efu_180lt_scr", "efu_1yr_scr", "efu_1yr_readmit",
  "efu_1yrlt_scr"
)

efu_raw <- list.files(efu_dir, pattern = "_09092026\\.csv$", full.names = TRUE) %>%
  set_names(~ str_remove(basename(.x), "_09092026\\.csv$")) %>%
  map_dfr(
    ~ read_csv(.x, col_types = cols(.default = col_character()), show_col_types = FALSE),
    .id = "efu_center"
  ) %>%
  mutate(
    id_key = norm_id(subjectid),
    across(all_of(efu_yesno), as.integer),
    # parse_number() keeps below-detection strings such as "<0.2" (-> 0.2)
    across(all_of(efu_numeric), readr::parse_number)
  ) %>%
  select(-subjectid)

master <- master %>% mutate(id_key = norm_id(subjectid))

efu_cohort <- master %>%
  inner_join(efu_raw, by = "id_key")
```

## Join QC

```{r efu-qc}
#| code-summary: "Show join QC"
qc <- tibble(
  check = c(
    "EFU rows (all site files)",
    "EFU rows matched to master",
    "EFU rows NOT in master",
    "Master patients (n)",
    "Master patients without EFU data",
    "Analytic cohort (EFU patients)"
  ),
  n = c(
    nrow(efu_raw),
    sum(efu_raw$id_key %in% master$id_key),
    sum(!efu_raw$id_key %in% master$id_key),
    nrow(master),
    sum(!master$id_key %in% efu_raw$id_key),
    nrow(efu_cohort)
  )
)
knitr::kable(qc, caption = "EFU vs master join")
```

```{r efu-qc-sites}
#| code-summary: "Show per-site join status"
bind_rows(
  efu_raw %>%
    mutate(status = if_else(id_key %in% master$id_key, "matched", "not in master")) %>%
    count(site_file = efu_center, status),
  master %>%
    filter(!id_key %in% efu_raw$id_key) %>%
    mutate(site_file = "(no EFU file)", status = "master only") %>%
    count(site_file, status)
) %>%
  pivot_wider(names_from = status, values_from = n, values_fill = 0) %>%
  knitr::kable(caption = "Rows per site file by join status (MCX = Mayo + MCM)")
```

# Derived Variables

```{r derived}
#| code-summary: "Show derived variables"
t1_numeric <- c(
  "age", "admit_meld_3", "day0_meld_3", "baseline_scr_new", "baseline_scr",
  "tbili_admit", "inr_admit", "sodium_admit", "albumin_admit", "map_day0_avg",
  "aclf_grade", "clif_c_score", "Discont_Day", "total_albumin",
  "totalalbumin_before_terli", "albvolume_before_terli"
)

efu_cohort <- efu_cohort %>%
  mutate(
    # Average of the four day-0 MAP readings. A recorded 0 mmHg is a
    # "not measured" placeholder, so it is set to NA before averaging.
    across(
      c(map_terli_day0_time0, map_terli_day0_time1,
        map_terli_day0_time2, map_terli_day0_time3),
      ~ na_if(as.numeric(.x), 0)
    ),
    map_day0_avg = rowMeans(
      across(c(map_terli_day0_time0, map_terli_day0_time1,
               map_terli_day0_time2, map_terli_day0_time3)),
      na.rm = TRUE
    ),
    map_day0_avg = if_else(is.nan(map_day0_avg), NA_real_, map_day0_avg),
    across(all_of(t1_numeric), as.numeric)
  )

tibble(
  `map_day0_avg` = c("non-missing", "readings used per patient (median)"),
  value = c(
    sum(!is.na(efu_cohort$map_day0_avg)),
    median(rowSums(!is.na(efu_cohort[, paste0("map_terli_day0_time", 0:3)])))
  )
) %>% knitr::kable()
```

## Listing status: index hospitalization vs EFU

`listed_transplant` (shared pipeline: `lt_eval == 2 | lteval_listingstatus == 1`)
captures listing during the index hospitalization; `efu_listingstatus`
captures listing at any point through 1 year.

```{r listing-crosstab}
#| code-summary: "Show cross-tabulation"
with(efu_cohort,
     table(listed_transplant = listed_transplant,
           efu_listingstatus = efu_listingstatus, useNA = "ifany")) %>%
  knitr::kable(caption = "Rows: listed_transplant; columns: efu_listingstatus (NA = missing)")
```

# Table 1 — Demographics by `efu_listingstatus`

Cohort: EFU patients with a non-missing `efu_listingstatus`. Continuous
variables are median [Q1, Q3] (Mann-Whitney U); categorical are n (%)
(chi-squared, or Fisher's exact when any cell < 5).

```{r table1}
#| code-summary: "Show Table 1 code"
t1_all_vars <- c(
  "age", "sex_male", "race", "cirrhosis_etiology",
  "he_previous_comp", "ascites_previous_comp", "refascites_previous_comp",
  "sbp_previous_comp", "nonsbp_previous_comp", "phtn_bleed_previous_comp",
  "hcc_previous_comp", "tips_previous_comp", "aki_previous_90days",
  "mido_preadmission", "oct_before_terli", "pressors_aki",
  "admit_meld_3", "day0_meld_3", "baseline_scr_new", "baseline_scr",
  "tbili_admit", "inr_admit", "sodium_admit", "albumin_admit",
  "map_day0_avg",
  "aclf_grade", "clif_c_score", "icu_admit",
  "criteria_1", "criteria_2", "criteria_3", "criteria_4",
  "terli_route_day0", "Discont_Day", "total_albumin",
  "albumin_before_terli", "totalalbumin_before_terli", "albvolume_before_terli",
  "hrs_responders", "hrs_responders_cat_2"
)
t1_cat_vars <- setdiff(t1_all_vars, t1_numeric)

t1_data <- efu_cohort %>% filter(!is.na(efu_listingstatus))

create_table_one(
  all_vars = t1_all_vars,
  cat_vars = t1_cat_vars,
  data     = t1_data,
  strata   = "efu_listingstatus",
  output   = "tables/table1_demographics_efu_listingstatus.csv"
)
```

[📥 Download CSV — Table 1](tables/table1_demographics_efu_listingstatus.csv)

# Table 2 — EFU Summary by `listed_transplant`

Cohort: all EFU patients. `efu_days*` fields are days from **index admission**.

```{r table2}
#| code-summary: "Show Table 2 code"
efu_vars <- c(
  "efu_readmit", "efu_daysreadmit", "efu_readmit_cirrhosiscomp", "efu_readmitaki",
  "efu_readmitscr", "efu_status", "efu_dayslastencounter", "efu_daysdod",
  "efu_rrt", "efu_daysrrtstart", "efu_listingstatus", "efu_dayslisting",
  "efu_listing", "efu_dayswlremoval", "efu_lt", "efu_dayslt", "efu_slkt",
  "efu_kal", "efu_dayskal", "efu_ltrrt", "efu_ltna", "efu_lttbili", "efu_ltscr",
  "efu_lt_albumin", "efu_lt_inr", "efu_90_scr", "efu_90_rrt", "efu_90_readmit",
  "efu_90lt_scr", "efu_90lt_rrt", "efu_180_scr", "efu_180_rrt", "efu_180_readmit",
  "efu_180lt_scr", "efu_180lt_rrt", "efu_1yr_scr", "efu_1yr_rrt", "efu_1yr_readmit",
  "efu_1yrlt_scr", "efu_1yrlt_rrt"
)
efu_cat_vars <- intersect(efu_vars, efu_yesno)

create_table_one(
  all_vars = efu_vars,
  cat_vars = efu_cat_vars,
  data     = efu_cohort,
  strata   = "listed_transplant",
  output   = "tables/table2_efu_listed_transplant.csv"
)
```

[📥 Download CSV — Table 2](tables/table2_efu_listed_transplant.csv)

# Table 3 — EFU Summary by `efu_listingstatus`

Cohort: EFU patients with a non-missing `efu_listingstatus`. Same rows as
Table 2 minus the stratifier.

```{r table3}
#| code-summary: "Show Table 3 code"
t3_vars <- setdiff(efu_vars, "efu_listingstatus")

create_table_one(
  all_vars = t3_vars,
  cat_vars = intersect(t3_vars, efu_yesno),
  data     = t1_data,
  strata   = "efu_listingstatus",
  output   = "tables/table3_efu_efu_listingstatus.csv"
)
```

[📥 Download CSV — Table 3](tables/table3_efu_efu_listingstatus.csv)

# Session Info

```{r session-info}
#| code-summary: "Show session info"
sessionInfo()
```
