---
title: "Liver Transplant — Analysis"
author: "Tianqi Ouyang"
date: last-modified
page-layout: full
format:
  html:
    echo: true
    code-fold: true
    toc: true
    toc-depth: 3
---

# Scope

Extended follow-up (EFU) of the HARMONY terlipressin cohort. The analytic
cohort is every patient in the locked master dataset
(`final_master_01282026.xlsx`, n = 243) who has an EFU record in the
site-submitted files under
`Extended Data Collection/Finalized Files/09SEP2026/` (13 site CSVs,
222 rows).

| Table | Cohort | Column (stratifier) | Rows |
|:--|:--|:--|:--|
| 1 — Demographics | EFU patients | `efu_listingstatus` (listed for LT at any point) | Baseline demographics, prior complications, admission / day-0 labs & scores, treatment characteristics, HRS response |
| 2 — EFU summary | EFU patients | `listed_transplant` (index-hospitalization listing, shared pipeline) | All EFU outcome fields |
| 3 — EFU summary | EFU patients | `efu_listingstatus` | All EFU outcome fields except the stratifier |

# Data Handling Notes

| Topic | Handling in this analysis |
|:--|:--|
| Study IDs | EFU files format Yale IDs as `YAL-00N`; the master uses `Yale-N`. Both sides are joined on a normalized key (upper-case prefix + integer). |
| Cohort size | 222 EFU rows → 221 analyzable patients: one CCF row has an ID that is not in the locked master; the 22 CSF master patients have no EFU file. The cohort n is computed, not hard-coded. |
| Day-0 MAP | `map_day0_avg` = mean of `map_terli_day0_time0..3` after recoding `0` (not measured) to `NA`. |
| Albumin during therapy | `albumintotal_terli_dayN` = grams given on day N; the requested `albumintotal_terli_dayX` row is the shared `total_albumin` (sum over days 0–13; nothing is recorded after `Discont_Day`). `alb_terli_dayN` is *serum* albumin and is not used. |
| Below-detection labs | One `efu_lttbili` value is a `<x.x` string; `readr::parse_number()` keeps the limit value. |
| Missing stratifier | One patient has no `efu_listingstatus`; Tables 1 and 3 use n = 220, Table 2 uses n = 221. |

```{r setup}
#| include: false
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(readxl)
library(tableone)

# Load `master` (n = 243) and the shared helpers. `output = tempfile()` keeps
# the purl()ed script out of the working directory.
source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  output = tempfile(fileext = ".R"),
  quiet  = TRUE
))

dir.create("tables", showWarnings = FALSE)
```

# Table Helper

Tables use the shared pipeline's `create_table_one()`:
`tableone::CreateTableOne()` → `print()` → `write.csv()`, with the same printed
matrix returned as a `knitr::kable()`, so the web table and the CSV are
identical. Conventions (tableone defaults, matching every earlier Table 1 CSV
in the registry): continuous variables as median [IQR] with the Kruskal-Wallis
rank-sum test (`test = nonnorm`; equivalent to Mann-Whitney for two groups);
categorical variables as n (%) with the chi-squared test (continuity-corrected
for 2 × 2). `Missing` is the percentage of the cohort with a missing value.

# EFU Data Import

Each site returned one CSV. All files share the same 42 columns
(`subjectid` + 41 EFU fields; see the *EFU Variables* page). The files
contain no dates — only REDCap-calculated day counts from the index
admission — so nothing patient-identifying is loaded beyond the study ID
used for the join.

```{r efu-import}
#| code-summary: "Show EFU import and ID normalization"
efu_dir <- "/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Extended Data Collection/Finalized Files/09SEP2026"

# Study IDs differ in formatting between REDCap exports (e.g. `YAL-003` in the
# EFU file vs `Yale-3` in the master). Normalize both sides to
# `<UPPERCASE PREFIX>-<integer>` before joining.
norm_id <- function(x) {
  x      <- str_trim(x)
  prefix <- toupper(str_extract(x, "^[A-Za-z]+"))
  prefix <- if_else(prefix == "YAL", "YALE", prefix)
  number <- as.integer(str_extract(x, "[0-9]+$"))
  paste0(prefix, "-", number)
}

# Field types from the EFU data dictionary
efu_yesno <- c(
  "codestatus", "efu_readmit", "efu_readmit_cirrhosiscomp", "efu_readmitaki",
  "efu_status", "efu_rrt", "efu_listingstatus", "efu_listing", "efu_lt",
  "efu_slkt", "efu_kal", "efu_ltrrt", "efu_90_rrt", "efu_90lt_rrt",
  "efu_180_rrt", "efu_180lt_rrt", "efu_1yr_rrt", "efu_1yrlt_rrt"
)
efu_numeric <- c(
  "efu_daysreadmit", "efu_readmitscr", "efu_dayslastencounter", "efu_daysdod",
  "efu_daysrrtstart", "efu_dayslisting", "efu_dayswlremoval", "efu_dayslt",
  "efu_dayskal", "efu_ltna", "efu_lttbili", "efu_ltscr", "efu_lt_albumin",
  "efu_lt_inr", "efu_90_scr", "efu_90_readmit", "efu_90lt_scr", "efu_180_scr",
  "efu_180_readmit", "efu_180lt_scr", "efu_1yr_scr", "efu_1yr_readmit",
  "efu_1yrlt_scr"
)

efu_raw <- list.files(efu_dir, pattern = "_09092026\\.csv$", full.names = TRUE) %>%
  set_names(~ str_remove(basename(.x), "_09092026\\.csv$")) %>%
  map_dfr(
    ~ read_csv(.x, col_types = cols(.default = col_character()), show_col_types = FALSE),
    .id = "efu_center"
  ) %>%
  mutate(
    id_key = norm_id(subjectid),
    across(all_of(efu_yesno), as.integer),
    # parse_number() keeps below-detection strings such as "<0.2" (-> 0.2)
    across(all_of(efu_numeric), readr::parse_number)
  ) %>%
  select(-subjectid)

master <- master %>% mutate(id_key = norm_id(subjectid))

efu_cohort <- master %>%
  inner_join(efu_raw, by = "id_key")
```

## Join QC

```{r efu-qc}
#| code-summary: "Show join QC"
qc <- tibble(
  check = c(
    "EFU rows (all site files)",
    "EFU rows matched to master",
    "EFU rows NOT in master",
    "Master patients (n)",
    "Master patients without EFU data",
    "Analytic cohort (EFU patients)"
  ),
  n = c(
    nrow(efu_raw),
    sum(efu_raw$id_key %in% master$id_key),
    sum(!efu_raw$id_key %in% master$id_key),
    nrow(master),
    sum(!master$id_key %in% efu_raw$id_key),
    nrow(efu_cohort)
  )
)
knitr::kable(qc, caption = "EFU vs master join")
```

```{r efu-qc-sites}
#| code-summary: "Show per-site join status"
bind_rows(
  efu_raw %>%
    mutate(status = if_else(id_key %in% master$id_key, "matched", "not in master")) %>%
    count(site_file = efu_center, status),
  master %>%
    filter(!id_key %in% efu_raw$id_key) %>%
    mutate(site_file = "(no EFU file)", status = "master only") %>%
    count(site_file, status)
) %>%
  pivot_wider(names_from = status, values_from = n, values_fill = 0) %>%
  knitr::kable(caption = "Rows per site file by join status (MCX = Mayo + MCM)")
```

# Derived Variables

```{r derived}
#| code-summary: "Show derived variables"
t1_numeric <- c(
  "age", "admit_meld_3", "day0_meld_3", "baseline_scr_new", "baseline_scr",
  "tbili_admit", "inr_admit", "sodium_admit", "albumin_admit", "map_day0_avg",
  "aclf_grade", "clif_c_score", "Discont_Day", "total_albumin",
  "totalalbumin_before_terli", "albvolume_before_terli"
)

efu_cohort <- efu_cohort %>%
  mutate(
    # Average of the four day-0 MAP readings. A recorded 0 mmHg is a
    # "not measured" placeholder, so it is set to NA before averaging.
    across(
      c(map_terli_day0_time0, map_terli_day0_time1,
        map_terli_day0_time2, map_terli_day0_time3),
      ~ na_if(as.numeric(.x), 0)
    ),
    map_day0_avg = rowMeans(
      across(c(map_terli_day0_time0, map_terli_day0_time1,
               map_terli_day0_time2, map_terli_day0_time3)),
      na.rm = TRUE
    ),
    map_day0_avg = if_else(is.nan(map_day0_avg), NA_real_, map_day0_avg),
    across(all_of(t1_numeric), as.numeric)
  )

tibble(
  `map_day0_avg` = c("non-missing", "readings used per patient (median)"),
  value = c(
    sum(!is.na(efu_cohort$map_day0_avg)),
    median(rowSums(!is.na(efu_cohort[, paste0("map_terli_day0_time", 0:3)])))
  )
) %>% knitr::kable()
```

## Listing status: index hospitalization vs EFU

`listed_transplant` (shared pipeline: `lt_eval == 2 | lteval_listingstatus == 1`)
captures listing during the index hospitalization; `efu_listingstatus`
captures listing at any point through 1 year.

```{r listing-crosstab}
#| code-summary: "Show cross-tabulation"
with(efu_cohort,
     table(listed_transplant = listed_transplant,
           efu_listingstatus = efu_listingstatus, useNA = "ifany")) %>%
  knitr::kable(caption = "Rows: listed_transplant; columns: efu_listingstatus (NA = missing)")
```

# Table 1 — Demographics by `efu_listingstatus`

Cohort: EFU patients with a non-missing `efu_listingstatus`. Continuous
variables are median [IQR] (Kruskal-Wallis / Mann-Whitney, `test = nonnorm`);
categorical are n (%) (chi-squared).

```{r table1}
#| code-summary: "Show Table 1 code"
t1_all_vars <- c(
  "age", "sex_male", "race", "cirrhosis_etiology",
  "he_previous_comp", "ascites_previous_comp", "refascites_previous_comp",
  "sbp_previous_comp", "nonsbp_previous_comp", "phtn_bleed_previous_comp",
  "hcc_previous_comp", "tips_previous_comp", "aki_previous_90days",
  "mido_preadmission", "oct_before_terli", "pressors_aki",
  "admit_meld_3", "day0_meld_3", "baseline_scr_new", "baseline_scr",
  "tbili_admit", "inr_admit", "sodium_admit", "albumin_admit",
  "map_day0_avg",
  "aclf_grade", "clif_c_score", "icu_admit",
  "criteria_1", "criteria_2", "criteria_3", "criteria_4",
  "terli_route_day0", "Discont_Day", "total_albumin",
  "albumin_before_terli", "totalalbumin_before_terli", "albvolume_before_terli",
  "hrs_responders", "hrs_responders_cat_2"
)
t1_cat_vars <- setdiff(t1_all_vars, t1_numeric)

t1_data <- efu_cohort %>% filter(!is.na(efu_listingstatus))

create_table_one(
  all_vars = t1_all_vars,
  cat_vars = t1_cat_vars,
  data     = t1_data,
  strata   = "efu_listingstatus",
  output   = "tables/table1_demographics_efu_listingstatus.csv",
  caption  = "Table 1 - Demographics by efu_listingstatus (0 = never listed, 1 = listed at any point)"
)
```

[📥 Download CSV — Table 1](tables/table1_demographics_efu_listingstatus.csv)

# Table 2 — EFU Summary by `listed_transplant`

Cohort: all EFU patients. `efu_days*` fields are days from **index admission**.

```{r table2}
#| code-summary: "Show Table 2 code"
efu_vars <- c(
  "efu_readmit", "efu_daysreadmit", "efu_readmit_cirrhosiscomp", "efu_readmitaki",
  "efu_readmitscr", "efu_status", "efu_dayslastencounter", "efu_daysdod",
  "efu_rrt", "efu_daysrrtstart", "efu_listingstatus", "efu_dayslisting",
  "efu_listing", "efu_dayswlremoval", "efu_lt", "efu_dayslt", "efu_slkt",
  "efu_kal", "efu_dayskal", "efu_ltrrt", "efu_ltna", "efu_lttbili", "efu_ltscr",
  "efu_lt_albumin", "efu_lt_inr", "efu_90_scr", "efu_90_rrt", "efu_90_readmit",
  "efu_90lt_scr", "efu_90lt_rrt", "efu_180_scr", "efu_180_rrt", "efu_180_readmit",
  "efu_180lt_scr", "efu_180lt_rrt", "efu_1yr_scr", "efu_1yr_rrt", "efu_1yr_readmit",
  "efu_1yrlt_scr", "efu_1yrlt_rrt"
)
efu_cat_vars <- intersect(efu_vars, efu_yesno)

create_table_one(
  all_vars = efu_vars,
  cat_vars = efu_cat_vars,
  data     = efu_cohort,
  strata   = "listed_transplant",
  output   = "tables/table2_efu_listed_transplant.csv",
  caption  = "Table 2 - EFU outcomes by listed_transplant (0 = not listed during index hospitalization, 1 = listed)"
)
```

[📥 Download CSV — Table 2](tables/table2_efu_listed_transplant.csv)

# Table 3 — EFU Summary by `efu_listingstatus`

Cohort: EFU patients with a non-missing `efu_listingstatus`. Same rows as
Table 2 minus the stratifier.

```{r table3}
#| code-summary: "Show Table 3 code"
t3_vars <- setdiff(efu_vars, "efu_listingstatus")

create_table_one(
  all_vars = t3_vars,
  cat_vars = intersect(t3_vars, efu_yesno),
  data     = t1_data,
  strata   = "efu_listingstatus",
  output   = "tables/table3_efu_efu_listingstatus.csv",
  caption  = "Table 3 - EFU outcomes by efu_listingstatus (0 = never listed, 1 = listed at any point)"
)
```

[📥 Download CSV — Table 3](tables/table3_efu_efu_listingstatus.csv)

# Session Info

```{r session-info}
#| code-summary: "Show session info"
sessionInfo()
```
